Hippocampal replay of experience at real world speeds

### 2021-06-02 Article.tex
Update on Overleaf.

% Figure 1
\begin{figure}
\includegraphics[width=0.75\linewidth]{figures/Figure1/Figure1_v7}
\caption{The model can capture different sequence dynamics on simulated data. \textbf{(A)} We construct a firing sequence of 19 simulated place cells that exhibits three different movement dynamics. For the first 60 ms, one cell fires repeatedly, representing one stationary location. For next 190 ms, the cells fire in sequence, representing a rapid continuous trajectory along the virtual track. For the last 30 ms, cells fire randomly, out of spatial order, representing an fragmented spatial sequence. \textbf{(B)} Like the standard decoder, the state space model uses estimates of cells' place fields from when the animal is moving and combines them with the observed spikes in (A) to compute the likelihood of position for each time step. \textbf{(C)} The prediction from the neural data is then combined with an explicit model of each movement dynamic, which determines how latent position can change based on the position in the previous time step. We show the probability of the next position bin for each movement dynamic model (color scale). Zero here represents the previous position. \textbf{(D)} The probability of remaining in a particular movement dynamic versus switching to another dynamic is modeled as having a high probability of remaining in a particular dynamic with a small probability of switching to one of the other dynamics at each time step. \textbf{(E)} The model uses the components in A-D over all time to decode the joint posterior probability of latent position and dynamic. This can be summarized by marginalizing over latent position (left panel) to get the probability of each dynamic over time. The shaded colors indicate the category of the speed of the trajectory at that time (Stat. = Stationary, S-C-M = Stationary-Continuous-Mixture, Cont. = Continuous, F-C-M = Fragmented-Continuous-Mixture, Frag. = Fragmented), which is determined from the probability. Marginalizing the posterior across dynamics also provides an estimate of latent position over time (right panel). Red dotted line in the right panel is the best fit line from the standard decoder using the Radon transform. \textbf{(F)} The probability of each dynamic depends heavily on the speed of the trajectory, as we show using a range of simulated spiking sequences each moving at a constant speed. Each dot corresponds to the average probability of that dynamic for a given constant speed sequence. We use a 0.8 threshold (dotted line) to classify each sequence based on the dynamic or dynamics which contribute maximally to the posterior (shaded colors).
}
\label{fig:1}
% Figure 1-Supplemental 1
\figsupp[The model is robust to change of the probability of persisting in the same dynamic for a wide range of plausible expected durations (25 ms to 150 ms)]{The model is robust to change of the probability of persisting in the same dynamic for a wide range of plausible expected durations (25 ms to 150 ms). \textbf{(A)} Each panel shows the probability of each dynamic on simulated data example from \FIG{1} with a different diagonal value---which governs the probability of remaining in that dynamic. The corresponding expected duration of staying in the dynamic is listed as duration. The off-diagonal values---the probability of switching to one of the other dynamics---are set to be equally likely with the remainder of the probability, as in \FIG{1}D. The diagonal increases from left to right, top to bottom, until the case where the diagonal is one and the off-diagonal is zero---i.e. the case where there is no probability of switching to another dynamic. Shaded regions correspond to the classification as in \FIG{1}F. \textbf{(B)} The probability of position over time for each diagonal value. Conventions the same as in (A).}{\includegraphics[width=1.0\linewidth]{figures/Figure1-supplemental1/Figure1_v3_supplemental1}}\label{figsupp:fig1sup1}
\end{figure}

@@

Decoding the spiking sequence requires specifying two elements: the data model---how the spikes relate to position---and the movement dynamic models---how the position can change over time in each movement dynamic. For the data model, our decoder is the same as the standard ("Bayesian") decoder \citep{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019}. We compute an estimate of how each cell's firing rate varies over position during movement (i.e. the place field, \FIG{1}B). This is used during decoding to compute the Poisson likelihood of position over time given the spiking sequence of interest. In contrast to the standard decoding approaches, we can use small time bins (in our case 2 ms vs. 20 ms or more) because we are able to take advantage of the prior placed on the dynamics by the state space model. This allows us to detect changes on smaller time scales than would be possible with the standard decoder and incorporate information about times when there is no spiking \cite{DengClusterlessDecodingPosition2015}. Further, because place estimates from spikes within a single bin are combined, our small time bins allow us to measure the spatial information conveyed by single spikes, rather than assuming that a downstream neuron would integrate spikes from multiple neurons on the timescale of 20 ms or more.

@@

This illustrates two key features of the model. First, as mentioned above, the fragmented dynamic gives us a way to directly identify times when the position representation is spatially incoherent, a higher resolution and less assumption-dependent alternative to the more commonly used non-parametric shuffle. Second, this approach allows the model to to capture a wide range of movement speeds for the latent position in contrast to the standard decoder line fit (red dashed line in \FIG{1}E, using the radon transform). The model is defined in terms of a mixture of movement dynamics, as summarized by the probability of each movement dynamic, and which dynamic or dynamics are dominant at a given moment is related to the temporal evolution of the underlying position representation. To demonstrate this, we applied the model to 10,000 simulated trajectories, each trajectory proceeding at a constant speed (\FIG{1}F) from 1 cm/s to 10,000 cm/s. From this, we can see that not only are there regions of speed that correspond to each of our three movement dynamics being highly probable (where we define highly probable to be greater than or equal to 0.8 probability), but there are also intermediate speeds where two of the dynamics exhibit relatively high probability; and where the sum of two of the dynamics' probabilities exceeds 0.8. In this manuscript, we will refer to these intermediate speeds as mixture dynamics. For example, when the stationary dynamic has a probability of 0.6 and the continuous has a probability of 0.4, we call this a stationary-continuous mixture (light blue, \FIG{1}F) and this indicates that the trajectory is moving slowly. Correspondingly, if the continuous dynamic has a probability of 0.5 and the fragmented dynamic has a probability of 0.4, then we would call this a fragmented-continuous-mixture and this indicates the trajectory is moving very quickly, but not as quickly as the fragmented dynamic dictates. In summary, we can characterize the speed or set of speeds that occur within a SWR based on the probability of each of the three movement dynamics over time. We further classify the probability of each movement dynamic as being part of one of five speed categories: stationary, stationary-continuous-mixtures, continuous, fragmented-continuous mixtures, and fragmented.

We note here that the choice of any particular threshold for classifying the movement dynamic of a SWR is arbitrary, and that the power of our approach lies in part on the ability to assign a probability for each dynamic or combinations of dynamics to each moment in time. Our goal in choosing 0.8 was to use a threshold that corresponds to highly probable events and to explicitly avoid the arbitrary choice of 0.95, as the 0.95 threshold often leads to the conclusion that only events that are statistically significant at the 0.05 level are meaningful in the brain. Nonetheless, we also verify that our central results hold with a higher threshold of 0.95.

@@

Using a 0.8 threshold, 89\% (23,071 of 25,844) of SWRs contain at least one of the three dynamics or dynamic mixtures. To ensure that this reflects the spatially tuned firing of the neurons, we trained the encoding model with positions resampled with replacement, a shuffling procedure which disrupts the relationship between spiking and position, for two recording sessions. We then decoded the same SWR events, containing the original spikes. Only 9\% of the SWRs are classified in the shuffled datasets, a value that is significantly less than that seen for the real data (p=0.02 for recording session 1, p=0.02 for recording session 2, \FIGSUPP[3]{fig3sup2}). This shows that our model does not impose movement dynamics in the absence of spatial information in the data.

% Figure 3
\begin{figure}
\includegraphics[width=0.8\linewidth]{figures/Figure3/Figure3_v5}
\caption{A-F. Six examples of SWRs with non-constant speed trajectories. Figure conventions are the same as in \FIG{2}. Filtered SWR ripple (150-250 Hz) trace from the tetrode with the maximum amplitude displayed above each example. \textbf{(A)} A SWR where the decoded position starts moving down the center arm away from the animal's position at the center well, slows down, and returns back. \textbf{(B)} A SWR where the decoded position persistently stays at the choice point (open circle) while the animal remains at the left well. \textbf{(C)} A SWR where the decoded position begins with stationary representation of the left well, then jumps to the middle of the right arm and proceeds up the right arm to the right well. \textbf{(D)} A SWR where the decoded position begins with stationary representation of the left well, jumps to the center arm, proceeds away from the center well, jumps to the right arm, proceeds back toward the center well, and then becomes fragmented. \textbf{(E)} A SWR where the decoded position begins in the left arm and persists at the end of the center arm. \textbf{(F)} A SWR where the decoded position starts in the left arm toward the choice point, jumps to the right arm and proceeds back toward the choice point. \textbf{(G)} Classification of SWRs from multiple animals and datasets. Each dot represents the percentage of SWRs for each day for that animal. A SWR is included in the numerator of the percentage if at any time it includes the classifications listed below the column. The denominator is listed in the x-axis label.
}
\label{fig:3}

@@

% Figure 3-Supplemental 2
\figsupp[Percentage of SWRs classified decreases on position shuffled data]{Comparison of percentage of SWRs classified (that is, a SWR containing at least one of the five classifications) on real vs. position shuffled data for two recording sessions from different animals. Red line represents the percent of SWR events classified in that recording session for real data. The histogram represents the distribution after 50 shuffles of the position data. Position data was shuffled by resampling with replacement from the set of all observed positions in that recording session, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure3-supplemental2/Figure3_v2-supplemental2}}\label{figsupp:fig3sup2}

% Figure 3-Supplemental 3
\figsupp[Percentage of spatially incoherent SWRs increases for run-swapped, circularly shuffled position data]{Comparison of percentage of spatially coherent and incoherent SWRs on real vs. run-swapped, circularly shuffled position. The red line represents the percent of spatially coherent or incoherent SWRs in that recording session for actual data. The histogram represents the distribution after 50 shuffles of the run from well to well as well as the circularly shuffled position (for each tetrode). This preserves local spatial correlations in the data but breaks the global spatial relationships between the spikes and the data.}{\includegraphics[width=0.8\linewidth]{figures/Figure3-supplemental3/Figure3_v1_supplemental3}}\label{figsupp:fig3sup3}

@@

Previous work focusing on spatially sequential replay reported that only a minority of events contain sequential spatial content \citep{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}. We therefore asked what fraction of classified events contain spatially coherent content, which we define as any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods and \FIG{1}F). We find that 96\% (22,170 of 23,071) of classified SWRs and 86\% of all SWRs include spatially coherent structure, and that this prevalence of spatially coherent structure is consistent across animals (\FIG{3}G). We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented or fragmented-continuous mixture dynamics. We find that only 14\% (3,295 of 23,071) of classified SWRs include any times with spatially incoherent structure (\FIG{3}G). To further validate this result, we performed a shuffling procedure that, in contrast to our previous shuffle, preserves the local correlations between spikes but reduces global spatial structure. This controls for dynamics that may have been induced by bursts or other local features of the spike train while randomizing the position to spike relationship. To do this, we randomized the order of runs (from one reward well to another) and then circularly permuted the resulting segments of data across all tetrodes uniformly. Using this shuffle, we found that there was still less spatially coherent and more spatially incoherent content compared to the real data (recording session \#1, spatially coherent: 99\% vs. 71\%, p=0.02, real vs. shuffled; spatially incoherent: 22\% vs. 61\%, p=0.02, real vs. shuffled; recording session \#2, spatially coherent: 95\% vs. 63\%, p=0.02, real vs. shuffled; spatially incoherent: 16\% vs. 79\%, p=0.02, real vs. shuffled; \FIGSUPP[3]{fig3sup3}).

@@

We repeated our classification analysis with a higher classification threshold of 0.95 to ensure that our result was not dependent on the threshold of 0.8. We find that, while this change slightly reduced the total fraction of classified SWRs (19,180 of 25,845 or 74\% of all SWRs), an even higher fraction of the classified SWRs (19,020 of 19,180 or 99\% classified SWRs) included spatially coherent content. Similarly, SWRs containing spatially incoherent content comprised a small fraction of the classified SWRs (484 of 19,180 or 3\% classified SWRs).

@@

% Figure 4-Supplemental 2
\figsupp[Shuffling the position data increases the average 95\% HPD region size for spatially coherent and incoherent classified times]{Average 95\% HPD region size increases for run-swapped, circularly shuffled position data compared to real data. Data from two recording sessions from different animals. Red line represents the average spatial position spanned by the highest posterior density on real data for all ripples. The histogram represents the distribution after 50 shuffles of the position data. The histogram represents the distribution after 50 shuffles of the run from well to well as well as the circularly shuffled position (for each tetrode). This preserves local correlations between spikes but reduces global spatial structure.}{\includegraphics[width=0.8\linewidth]{figures/Figure4-supplemental2/Figure4_v1-supplemental2}}\label{figsupp:fig4sup2}

@@

% Figure 5-Supplemental 2
\figsupp[Distribution of dynamics is insensitive to parameters]{Conventions are the same as in \FIG{5}.}{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental2/Figure5_v1_supplemental2}}\label{figsupp:fig5sup2}

@@

The second most prevalent classification was exclusively stationary events (4,858 of 21,433 or 23\% classified SWRs; 47\% of SWRs with stationary dynamics were entirely stationary). Unlike the stationary-continuous mixtures, most of these events represented a location close to the animal's actual position (\FIG{5}C). There were, however, a number of stationary events that represented positions far from the animal. We set a threshold of 30 cm distance from the animal to identify non-local stationary events and found such content in \texttildelow7\% of classified SWRs (1,437 of 21,433 classified SWRs). Of these, 46\% (664 of 1,437 stationary SWRs) were stationary throughout the duration of the SWR. These non-local stationary events most commonly represented reward wells or choice points (\FIG{5}E), consistent with \citep{YuDistincthippocampalcorticalmemory2017}, but a small portion of them occurred at other locations of the track. This suggests that these representations can be flexibly deployed to represent discrete locations.

@@

\subsection*{Classification of Dynamics}
We used a threshold of 0.8 to classify the probability of each state $Pr(I_{k} \mid O_{1:T})$ into 5 categories. Time periods during sharp wave ripples are labeled as stationary, continuous, or fragmented when the probability of each individual state is above 0.8. Time periods are labeled as stationary-continuous-mixture or fragmented-continuous-mixture when the sum of stationary and continuous or fragmented and continuous are above 0.8, respectively. Time periods where none of these criterion are met are considered unclassified.

>>>>

% Figure 1
\begin{figure}
\includegraphics[width=0.75\linewidth]{figures/Figure1/Figure1_v7}
\caption{The model can capture different sequence dynamics on simulated data. \textbf{(A)} We construct a firing sequence of 19 simulated place cells that exhibits three different movement dynamics. For the first 60 ms, one cell fires repeatedly, representing one stationary location. For next 190 ms, the cells fire in sequence, representing a rapid continuous trajectory along the virtual track. For the last 30 ms, cells fire randomly, out of spatial order, representing an fragmented spatial sequence. \textbf{(B)} Like the standard decoder, the state space model uses estimates of cells' place fields from when the animal is moving and combines them with the observed spikes in (A) to compute the likelihood of position for each time step. \textbf{(C)} The prediction from the neural data is then combined with an explicit model of each movement dynamic, which determines how latent position can change based on the position in the previous time step. We show the probability of the next position bin for each movement dynamic model (color scale). Zero here represents the previous position. \textbf{(D)} The probability of remaining in a particular movement dynamic versus switching to another dynamic is modeled as having a high probability of remaining in a particular dynamic with a small probability of switching to one of the other dynamics at each time step. \textbf{(E)} The model uses the components in A-D over all time to decode the joint posterior probability of latent position and dynamic. This can be summarized by marginalizing over latent position (left panel) to get the probability of each dynamic over time. The shaded colors indicate the category of the speed of the trajectory at that time (Stat. = Stationary, S-C-M = Stationary-Continuous-Mixture, Cont. = Continuous, F-C-M = Fragmented-Continuous-Mixture, Frag. = Fragmented), which is determined from the probability. Marginalizing the posterior across dynamics also provides an estimate of latent position over time (right panel). Red dotted line in the right panel is the best fit line from the standard decoder using the Radon transform. \textbf{(F)} The probability of each dynamic depends heavily on the speed of the trajectory, as we show using a range of simulated spiking sequences each moving at a constant speed. Each dot corresponds to the average probability of that dynamic for a given constant speed sequence. We use a 0.80 threshold (dotted line) to classify each sequence based on the dynamic or dynamics which contribute maximally to the posterior (shaded colors).
}
\label{fig:1}
% Figure 1-Supplemental 1
\figsupp[The model is robust to change of the probability of persisting in the same dynamic for a wide range of plausible expected durations (25 ms to 150 ms)]{The model is robust to change of the probability of persisting in the same dynamic for a wide range of plausible expected durations (25 ms to 150 ms). \textbf{(A)} Each panel shows the probability of each dynamic on simulated data example from \FIG{1} with a different diagonal value---which governs the probability of remaining in that dynamic. The corresponding expected duration of staying in the dynamic is listed as duration. The off-diagonal values---the probability of switching to one of the other dynamics---are set to be equally likely with the remainder of the probability, as in \FIG{1}D. The diagonal increases from left to right, top to bottom, until the case where the diagonal is one and the off-diagonal is zero---i.e. the case where there is no probability of switching to another dynamic. Shaded regions correspond to the classification as in \FIG{1}F. \textbf{(B)} The probability of position over time for each diagonal value. Conventions the same as in (A).}{\includegraphics[width=1.0\linewidth]{figures/Figure1-supplemental1/Figure1_v3_supplemental1}}\label{figsupp:fig1sup1}
\end{figure}

@@

Decoding the spiking sequence requires specifying two elements: the data model---how the spikes relate to position---and the movement dynamic models---how the position can change over time in each movement dynamic. For the data model, our decoder is the same as the standard ("Bayesian") decoder \citep{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019}. We compute an estimate of how each cell's firing rate varies over position during movement (i.e. the place field, \FIG{1}B). This is used during decoding to compute the Poisson likelihood of position over time given the spiking sequence of interest. In contrast to the standard decoding approaches, we can use small time bins (in our case 2 ms vs. 20 ms or more) because we are able to take advantage of the prior placed on the dynamics by the state space model. This allows us to detect changes on smaller time scales than would be possible with the standard decoder and incorporate information about times when there is no spiking \citep{DengClusterlessDecodingPosition2015}. Further, because place estimates from spikes within a single bin are combined, our small time bins allow us to measure the spatial information conveyed by single spikes, rather than assuming that a downstream neuron would integrate spikes from multiple neurons on the timescale of 20 ms or more.

@@

This illustrates two key features of the model. First, as mentioned above, the fragmented dynamic gives us a way to directly identify times when the position representation is spatially incoherent, a higher resolution and less assumption-dependent alternative to the more commonly used non-parametric shuffle. Second, this approach allows the model to to capture a wide range of movement speeds for the latent position in contrast to the standard decoder line fit (red dashed line in \FIG{1}E, using the radon transform). The model is defined in terms of a mixture of movement dynamics, as summarized by the probability of each movement dynamic, and which dynamic or dynamics are dominant at a given moment is related to the temporal evolution of the underlying position representation. To demonstrate this, we applied the model to 10,000 simulated trajectories, each trajectory proceeding at a constant speed (\FIG{1}F) from 1 cm/s to 10,000 cm/s. From this, we can see that not only are there regions of speed that correspond to each of our three movement dynamics being highly probable (where we define highly probable to be greater than or equal to 0.80 probability), but there are also intermediate speeds where two of the dynamics exhibit relatively high probability; and where the sum of two of the dynamics' probabilities exceeds 0.80. In this manuscript, we will refer to these intermediate speeds as mixture dynamics. For example, when the stationary dynamic has a probability of 0.6 and the continuous has a probability of 0.4, we call this a stationary-continuous mixture (light blue, \FIG{1}F) and this indicates that the trajectory is moving slowly. Correspondingly, if the continuous dynamic has a probability of 0.5 and the fragmented dynamic has a probability of 0.4, then we would call this a fragmented-continuous-mixture and this indicates the trajectory is moving very quickly, but not as quickly as the fragmented dynamic dictates. In summary, we can characterize the speed or set of speeds that occur within a SWR based on the probability of each of the three movement dynamics over time. We further classify the probability of each movement dynamic as being part of one of five speed categories: stationary, stationary-continuous-mixtures, continuous, fragmented-continuous mixtures, and fragmented.

We note here that the choice of any particular threshold for classifying the movement dynamic of a SWR is arbitrary, and that the power of our approach lies in part on the ability to assign a probability for each dynamic or combinations of dynamics to each moment in time. Our goal in choosing 0.80 was to use a threshold that corresponds to highly probable events and to explicitly avoid the arbitrary choice of 0.95, as the 0.95 threshold often leads to the conclusion that only events that are statistically significant at the 0.05 level are meaningful in the brain. Nonetheless, we also verify that our central results hold with a higher threshold of 0.95.

@@

Using a 0.80 threshold, 89\% (23,071 of 25,844) of SWRs contain at least one of the three dynamics or dynamic mixtures. To ensure that this reflects the spatially tuned firing of the neurons, we trained the encoding model with positions resampled with replacement, a shuffling procedure which disrupts the relationship between spiking and position, for two recording sessions. We then decoded the same SWR events, containing the original spikes. Only 9\% of the SWRs are classified in the shuffled datasets, a value that is significantly less than that seen for the real data (p=0.02 for recording session 1, p=0.02 for recording session 2, \FIGSUPP[3]{fig3sup2}). This shows that our model does not impose movement dynamics in the absence of spatial information in the data.

% Figure 3
\begin{figure}
\includegraphics[width=0.8\linewidth]{figures/Figure3/Figure3_v5}
\caption{A-F. Examples of SWRs with non-constant speed trajectories. Figure conventions are the same as in \FIG{2}. Filtered SWR ripple (150-250 Hz) trace from the tetrode with the maximum amplitude displayed above each example. \textbf{(A)} A SWR where the decoded position starts moving down the center arm away from the animal's position at the center well, slows down, and returns back. \textbf{(B)} A SWR where the decoded position persistently stays at the choice point (open circle) while the animal remains at the left well. \textbf{(C)} A SWR where the decoded position begins with stationary representation of the left well, then jumps to the middle of the right arm and proceeds up the right arm to the right well. \textbf{(D)} A SWR where the decoded position begins with stationary representation of the left well, jumps to the center arm, proceeds away from the center well, jumps to the right arm, proceeds back toward the center well, and then becomes fragmented. \textbf{(E)} A SWR where the decoded position begins in the left arm and persists at the end of the center arm. \textbf{(F)} A SWR where the decoded position starts in the left arm toward the choice point, jumps to the right arm and proceeds back toward the choice point. \textbf{(G)} Classification of SWRs from multiple animals and datasets. Each dot represents the percentage of SWRs for each day for that animal. A SWR is included in the numerator of the percentage if at any time it includes the classifications listed below the column. The denominator is listed in the x-axis label.
}
\label{fig:3}

@@

% Figure 3-Supplemental 2
\figsupp[Shuffling the position data with replacement decreases the percent of SWRs classified]{Comparison of percentage of SWRs classified (that is, a SWR containing at least one of the five classifications) on real vs. position shuffled data for two recording sessions from different animals. Red line represents the percent of SWR events classified in that recording session for real data. The histogram represents the distribution after 50 shuffles of the position data. Position data was shuffled by resampling with replacement from the set of all observed positions in that recording session, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure3-supplemental2/Figure3_v2-supplemental2}}\label{figsupp:fig3sup2}

% Figure 3-Supplemental 3
\figsupp[Shuffling the data by swapping the runs and circularly permuting the position increases the percentage of spatially incoherent SWRs and decreases the spatially coherent SWRs]{Shuffling the data by swapping the runs and circularly permuting the position increases the percentage of spatially incoherent SWRs and decreases the spatially coherent SWRs. The red line represents the percent of spatially coherent or incoherent SWRs in that recording session for actual data. The histogram represents the distribution after 50 shuffles of the run from well to well as well as the circularly shuffled position (for each tetrode). This preserves local spatial correlations in the data but breaks the global spatial relationships between the spikes and the data.}{\includegraphics[width=0.8\linewidth]{figures/Figure3-supplemental3/Figure3_v1_supplemental3}}\label{figsupp:fig3sup3}

@@

Previous work focusing on spatially sequential replay reported that only a minority of events contain sequential spatial content \citep{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}. We therefore asked what fraction of classified events contain spatially coherent content, which we define as any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods and \FIG{1}F). We find that 96\% (22,170 of 23,071) of classified SWRs and 86\% (22,170 of 25,844) of all SWRs include spatially coherent structure, and that this prevalence of spatially coherent structure is consistent across animals (\FIG{3}G). We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented or fragmented-continuous mixture dynamics. We find that only 14\% (3,295 of 23,071) of classified SWRs include any times with spatially incoherent structure (\FIG{3}G). To further validate this result, we performed a shuffling procedure that, in contrast to our previous shuffle, preserves the local correlations between spikes but reduces global spatial structure. This controls for dynamics that may have been induced by bursts or other local features of the spike train while randomizing the position to spike relationship. To do this, we randomized the order of runs (from one reward well to another) and then circularly permuted the resulting segments of data across all tetrodes uniformly. Using this shuffle, we found that there was still less spatially coherent and more spatially incoherent content compared to the real data (recording session \#1, spatially coherent: 99\% vs. 71\%, p=0.02, real vs. shuffled; spatially incoherent: 22\% vs. 61\%, p=0.02, real vs. shuffled; recording session \#2, spatially coherent: 95\% vs. 63\%, p=0.02, real vs. shuffled; spatially incoherent: 16\% vs. 79\%, p=0.02, real vs. shuffled; \FIGSUPP[3]{fig3sup3}).

@@

We repeated our classification analysis with a higher classification threshold of 0.95 to ensure that our result was not dependent on the threshold of 0.80. We find that, while this change slightly reduced the total fraction of classified SWRs (19,478 of 26,159 or 74\% of all SWRs), an even higher fraction of the classified SWRs (19,317 of 19,478 or 99\% classified SWRs) included spatially coherent content. Similarly, SWRs containing spatially incoherent content comprised a small fraction of the classified SWRs (490 of 19,478 or 3\% classified SWRs).

@@

% Figure 4-Supplemental 2
\figsupp[Shuffling the position data by swapping the runs and circularly permuting the data increases the average 95\% HPD region size for spatially coherent and incoherent classified times]{Shuffling the position data by swapping the runs and circularly permuting the data increases the average 95\% HPD region size for spatially coherent and incoherent classified times. Data from two recording sessions from different animals. Red line represents the average spatial position spanned by the highest posterior density on real data for all ripples. The histogram represents the distribution after 50 shuffles of the position data. The histogram represents the distribution after 50 shuffles of the run from well to well as well as the circularly shuffled position (for each tetrode). This preserves local correlations between spikes but reduces global spatial structure.}{\includegraphics[width=0.8\linewidth]{figures/Figure4-supplemental2/Figure4_v1-supplemental2}}\label{figsupp:fig4sup2}

@@

% Figure 5-Supplemental 2
\figsupp[Control analyses for distribution of dynamics]{Control analyses for distribution of dynamics. Conventions are the same as in \FIG{5}.}{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental2/Figure5_v1_supplemental2}}\label{figsupp:fig5sup2}

@@

The second most prevalent classification was exclusively stationary events (4,858 of 21,433 or 23\% classified SWRs; 47\% of SWRs with stationary dynamics were entirely stationary for the entire duration of the SWR). Unlike the stationary-continuous mixtures, most of these events represented a location close to the animal's actual position (\FIG{5}C). There were, however, a number of stationary events that represented positions far from the animal. We set a threshold of 30 cm distance from the animal to identify non-local stationary events and found such content in \texttildelow7\% of classified SWRs (1,437 of 21,433 classified SWRs). Of these, 46\% (664 of 1,437 stationary SWRs) were stationary throughout the duration of the SWR. These non-local stationary events most commonly represented reward wells or choice points (\FIG{5}E), consistent with \citep{YuDistincthippocampalcorticalmemory2017}, but a small portion of them occurred at other locations of the track. This suggests that these representations can be flexibly deployed to represent discrete locations.

@@

\subsection*{Classification of Dynamics}
We used a threshold of 0.80 to classify the probability of each state $Pr(I_{k} \mid O_{1:T})$ into 5 categories. Time periods during sharp wave ripples are labeled as stationary, continuous, or fragmented when the probability of each individual state is above 0.80. Time periods are labeled as stationary-continuous-mixture or fragmented-continuous-mixture when the sum of stationary and continuous or fragmented and continuous are above 0.80, respectively. Time periods where none of these criterion are met are considered unclassified.
